\chapter{Literature Review}
\label{cha:LitReview}

\section{Preliminaries}
\label{sec:LitReviewPrelim}

\section{Why Map the Internet?}
\label{sec:LitReviewWhyMap}

The main aim of this project is to embed a representation of the internet into hyperbolic space and produce a visualisation of this embedding. In order to fully understand why this is useful, we must ask the question: why do we need a map of the internet? 

Each node on the internet is capable of detecting which nodes it is connected to, however, it is not possible without special software to see beyond this. As each node only has local data it is not possible for them to form a complete model of the internet themselves and they must cooperate with other nodes to do this. Of course, this means that in order to gain a complete model of the internet it would be necessary to communicate with almost all of the nodes currently connected to it. Having a pre-computed map of the internet allows a node to know more about the global internet it is connected to without having to query almost every other node. 

Understanding the physical structure of the internet is something that network engineers have to deal with every day, however it can often be difficult to pinpoint problems in networks. For example, in standard Ethernet networks, the presence of a loop can often create a broadcast storm which can render entire parts of a network unusable due to packet flooding. If an accurate map of the network is known in advance then it is possible for an engineer to locate the fault more quickly as they can compare this map to the current state of the network and check to see if any new connections have been added which may be causing the loop. 

While maps may assist engineers on small scale problems they can also assist in diagnosing problems which affect the internet on a more global scale. For example, despite the extreme interconnectedness of the internet it is still possible for a specific node or link to become congested. Congestion of a node or link can cause severe packet loss making data transmission unreliable across the affected area. Having an accurate map of the internet would allow for quick diagnosis of the fault, as well as pinpointing the cause. Though it may have been possible without a map to easily find the affected nodes or links it would be far more difficult to find the cause of the issue. With a map, it is possible to identify features such as a large number of new nodes which may have caused increased data transmission across the affected area. Additionally, a map of the internet with a geographical element may assist in locating pairs of nodes which could receive additional links or additional bandwidth on existing links.

\section{The World Wide Web}
\label{sec:LitReviewWWW}
There is previous work \cite{munzner_visualizing_1995}, which consisted of visualising the structure of the World Wide Web (WWW) in hyperbolic space. Though this may seem similar to the aim of this project there are key differences between the internet and the WWW which creates sufficient difference. In order to appreciate this difference it is necessary to understand what the WWW is. 

While the internet is a physical concept, the WWW is not. The WWW represents the data that users of the internet can view, specifically it contains text data formatted using the Hypertext Mark-up Language (HTML) standard which allows users to view as well as navigate this data. This is done using a type of application known as a \textit{web browser}. Web browsers allow users to navigate to web addresses (\textit{www.google.co.uk} for example) and view information stored at these addresses. The user views web pages which can also contain links to other web pages, users can follow these links in their browser to visit these other pages. 

If we imagine each web page as a node in a graph and the links between pages as edges then it possible to construct a graph which represents the WWW or a subset thereof. Recall, however that the internet is, at a basic level, a graph too. While these two graphs have similarities, as stated earlier, they are fundamentally different. While the graph that represents the internet contains at most a single node per physical device*, it is possible for the graph that represents the WWW to contain multiple nodes per physical device. This is explained by having a node in the internet graph which hosts multiple pages in the WWW, which is extremely common. Due to this one to many mapping between physical internet nodes and WWW nodes the graphs of the internet and WWW are very different. 

A further difference between the two concepts is that links in the WWW have a direction whereas internet links do not**. It would be necessary for any graph representing the WWW to be able to represent directional links between nodes.  The internet graph does not need such complexity and can be represented with only bi-directional links. 

\section{Hyperbolic Space}
\label{sec:LitReviewHyperbolicSpace}

Hyperbolic space (or, more broadly, hyperbolic geometry) is by no means a recent concept and was postulated by Carl Friedrich Gauss in 1824 \cite{ratcliffe_foundations_2006}. In order to understand what hyperbolic geometry is, it is first necessary to review what is often described as standard: Euclidean geometry.

Euclidean geometry (postulated by Euclid in around 300 B.C. \cite{ratcliffe_foundations_2006}) is based upon five postulates, which Euclid described as follows \cite{ratcliffe_foundations_2006}:
\begin{enumerate}
\item \textit{A straight line may be drawn from any point to any other point.}
\item \textit{A finite straight line may be extended continuously in a straight line.}
\item \textit{A circle may be drawn with any centre and any radius.}
\item \textit{All right angles are equal.}
\item \textit{If a straight line falling on two straight lines makes the interior angles on the same side less than two right angles, the two straight lines, if extended infinitely, meet on the side on which the angles are less than two right angles.}
\end{enumerate}

The first four of these postulates seem both simple and natural, however the fifth is more involved and it is not immediately apparent what is being described by it. Euclid's fifth postulate effectively states that given a straight line and a point, there is only \textbf{one} straight line which passes through the point and is parallel to the original straight line. 

When considering the first four postulates it is easy to see that they should simply be accepted without proof due to their simplicity. However, the fifth postulate cannot be so easily accepted. A solution to this is to attempt to prove the fifth postulate using the others, something which was attempted for over two thousand years \cite{ratcliffe_foundations_2006} but was not possible. The only remaining hypothesis is that the fifth postulate is not required for a definition of geometry!

This seems absurd, as it is relatively simple to demonstrate in two dimensional Euclidean geometry that the fifth postulate holds, using only a piece of paper and a pencil. However, this does not prove that the postulate holds in all cases, only in this one instance. In order to refute the postulate, it is necessary cast out this rigid view of a a flat plane and to consider other geometries in which to work. 

If, instead of a plane, all geometry took place on the surface of a sphere then very different behaviour is observed. Of course, the surface of the sphere is a two dimensional plane but with one key difference: it is finite. On this finite plane a straight line is any which is a great circle of the sphere. In this case, it is clear that any straight line must have \textbf{zero} straight lines which are parallel to it, as all great circles must meet at exactly two points. 